---
title: "PA1_template"
author: "Choong-Hoon Hyun"
date: "5/7/2022"
output: html_document
---
# Peer-graded Assignment: Reproducible Research Week 2 Course Proejct 1

## Loading and preprocessing the data 

### Download file, load the dataset to R, and process/transforam the data into a format suitable for the analysis
```{r Loading and Preprocessing the data into a format sutable for the analysis, echo=TRUE}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
f <- file.path(getwd(), "activity.zip")
download.file(fileUrl, f, curl = "method")
unzip(zipfile = "activity.zip")

activity_monitoring_data <- read.csv("activity.csv")
activity_monitoring_data$date <- as.Date(activity_monitoring_data$date, "%Y-%m-%d")

```

## What is mean total number of steps taken per day?

###  1. Calculate the total number of steps taken per day
#### Load ggplot2
```{r loading ggplot2, echo=TRUE}
library(ggplot2)
```

```{r Make dataset for the total number of steps taken per day, echo=TRUE}
steps_per_day <- split(activity_monitoring_data, activity_monitoring_data$date)
total_steps <- sapply(steps_per_day, function(x) {
  sum(x[, "steps"])
})
total_steps_df <- as.data.frame(total_steps)
head(total_steps_df)
```
### 2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
#### Plot a histogram of the total number of steps taken each day
```{r Histogram of the total number of steps taken each day, echo=TRUE}
ggplot(total_steps_df, aes(total_steps)) + geom_histogram(binwidth = 1000, fill = "white", col = "blue") +
  labs(x= "Steps", y = "Frequency", title = "Total Number of Steps Taken each Day (Histogram)")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
#### Mean of the total number of steps taken per day
```{r Mean of the total number of steps taken per day, echo=TRUE}
mean(total_steps_df$total_steps, na.rm = TRUE)
```
#### Median of the total number of steps taken per day
```{r Median of the total number of steps taken per day, echo=TRUE}
median(total_steps_df$total_steps, na.rm = TRUE)
```
